Estimation of Intracortical Connectivity in a Dynamical Neural Field Model

### 2010-03-17 ltx/BrainIDE.tex
changes to equation (4) and (11)

where $f$ is spiking rate of populations $r'$, $w$ is the spatial connectivity kernel, and $\Omega$ is the spatial domain, representing a cortical sheet or surface. The term $I\left(r,t\right)$ accounts for spontaneous activity of the neural populations. The connectivity kernel is a Mexican hat shape described by the sum of weighted Gaussian functions
\begin{eqnarray}\label{Connectivity Kernel}
	w\left(r,r'\right) &=& \theta^T \psi\left(r,r'\right) \\
	\psi\left(r,r'\right) &=&\exp{\left(-\frac{1}{2}(r'-r)^T\Sigma_w (r'-r)\right)}
\end{eqnarray}
where
\begin{equation}
	\theta=\begin{bmatrix} \theta_{1}\\\theta_{2}\\\theta_{3}\end{bmatrix},\quad 		\Sigma_w=\begin{bmatrix}\Sigma_{w1}\\\Sigma_{w2}\\ \Sigma_{w3}\end{bmatrix}.
\end{equation}
\begin{todo}
	{Parham ... how are you writing this? DF: I have tried to make it better, but Parham can you check.}
\end{todo}The average firing rate of the presynaptic neurons is related to the postsynaptic membrane potential by the sigmoidal activation function
\begin{equation}\label{ActivationFunction}
	f\left( v\left( r', \bar t \right) \right) = \frac{\nu _0\left(r\right)}{1 + \exp \left( \beta \left( v_0\left( r \right) - v\left(r',t\right) \right) \right)}.
\end{equation}
Even though we will model $v_0$ and $\nu_0$ as being homogeneous across the cortical sheet we will express it as being spatially dependant allowing for generality at this point. By combining Eqs.~\ref{SpikesToPotential} and \ref{RateBasedInteractions} we get the spatiotemporal model
\begin{equation}\label{FullDoubleIntModel}
	v\left(r,t\right) = \int_{-\infty}^t h\left(t - t'\right) \left(\int_\Omega   w\left(r,r'\right) f\left( v\left( r',t \right)\right)dr' + I(r,t) \right)dt'
\end{equation}
% The rearrange the model into an integro-differential we take the Laplace transform of~\ref{FullDoubleIntModel} giving
% \begin{equation}
% 	V(s) = H(s)G(s).
% \end{equation}
% Since $H(s) = (s+\alpha)^{-1}$ we can write
% \begin{equation}
% 	(s+\alpha)V(s) = G(s).
% \end{equation}
To arrive at the final form of the model we shall state the synaptic response kernel as a Green's function
\begin{equation}\label{GreensFuncDef}
	Lh\left( t \right) = \delta \left( t \right),
\end{equation}
where $L=\frac{d}{dt} + \alpha$ is a temporal differential operator and $\delta(t)$ is the Dirac-delta function giving 
\begin{equation}\label{FinalForm1}
	\frac{dv\left( r,t \right)}{dt} + \alpha v\left( r,t \right) = \int_\Omega  {w\left( r,r' \right)f\left( {v\left( r',t \right)} \right)dr'} + I\left(r,t\right).
\end{equation}

\section{Generating Data}
To generate iEEG/LFP data we proceeded with the assumption of spatial homogeneity for the connectivity and activation function parameters, implying $w(r,r') = w\left(|r-r'|\right)$, $\nu_0(r) = \nu_0$ and $v_0(r) = v_0$. In addition we assume the synaptic transmission delay is negligible. The model was discretized in time  (see Appendix~\ref{Time Discretization}) and a stochastic term was added to drive the system and account for uncertainty in the model giving
\begin{equation}\label{DiscreteTimeModel}
	v_{t+1}\left(r\right) = \lambda v_t\left(r\right) + \delta \int_\Omega { w\left(|r-r'|\right) f\left(v_t\left(r'\right)\right) dr'} + \delta I_t\left(r\right) + e_t\left(r\right)
\end{equation}
where $\delta$ is the time step, $\lambda = 1-\alpha\delta$ and $e_t(r) \sim \mathcal{GP}(0,\Sigma)$. The IDE formulation in Eq.~\ref{DiscreteTimeModel} was solved using a numerical method described in Appendix~\ref{Space Discretization}. To generate iEEG/LFP data we used the output function 
\begin{equation}
	y_t = \int_{\Omega}{m\left(|r_n-r'|\right)v_t\left(r'\right)dr'} + \varepsilon_t
\end{equation}
where $r_n$ defines the location of the sensors in the field where $n=1,...,N$ indexes the sensors and $\varepsilon_t \sim \mathcal{N}\left(0,\Sigma_y\right)$. The output kernel $m(|r-r'|)$ governs the geometry of pick-up range of sensors where
\begin{equation}
	m\left(|r-r'|\right) = \exp{\left(-\frac{1}{2}(r'-r)^T\Sigma_m (r'-r)\right)}.
\end{equation}

>>>>

where $f$ is spiking rate of populations $r'$, $w$ is the spatial connectivity kernel, and $\Omega$ is the spatial domain, representing a cortical sheet or surface. The term $I\left(r,t\right)$ accounts for spontaneous activity of the neural populations. The connectivity kernel is a Mexican hat shape described by the sum of weighted Gaussian functions placed at the origin.
\begin{eqnarray}\label{Connectivity Kernel}
	w\left(r,r'\right) =&& \\
&&\theta_1\exp{\left(-\frac{(r-r')^\top(r-r')}{\sigma_1^2}\right)}+\theta_2\exp{\left(-\frac{(r-r')^\top(r-r')}{\sigma_2^2}\right)}+\theta_3\exp{\left(-\frac{(r-r')^\top(r-r')}{\sigma_3^2}\right)}\nonumber
\end{eqnarray}
where $\sigma_i$ is the scale parameter.
% where
% \begin{equation}
% 	\theta=\begin{bmatrix} \theta_{1}\\\theta_{2}\\\theta_{3}\end{bmatrix},\quad 		\Sigma_w=\begin{bmatrix}\Sigma_{w1}\\\Sigma_{w2}\\ \Sigma_{w3}\end{bmatrix}.
% \end{equation}
\begin{todo}
	{Parham ... how are you writing this? DF: I have tried to make it better, but Parham can you check.Parham: I have changed it but I think it's better to include it in the simulation section}
\end{todo}The average firing rate of the presynaptic neurons is related to the postsynaptic membrane potential by the sigmoidal activation function
\begin{equation}\label{ActivationFunction}
	f\left( v\left( r', \bar t \right) \right) = \frac{\nu _0\left(r\right)}{1 + \exp \left( \beta \left( v_0\left( r \right) - v\left(r',t\right) \right) \right)}.
\end{equation}
Even though we will model $v_0$ and $\nu_0$ as being homogeneous across the cortical sheet we will express it as being spatially dependant allowing for generality at this point. By combining Eqs.~\ref{SpikesToPotential} and \ref{RateBasedInteractions} we get the spatiotemporal model
\begin{equation}\label{FullDoubleIntModel}
	v\left(r,t\right) = \int_{-\infty}^t h\left(t - t'\right) \left(\int_\Omega   w\left(r,r'\right) f\left( v\left( r',t \right)\right)dr' + I(r,t) \right)dt'
\end{equation}
% The rearrange the model into an integro-differential we take the Laplace transform of~\ref{FullDoubleIntModel} giving
% \begin{equation}
% 	V(s) = H(s)G(s).
% \end{equation}
% Since $H(s) = (s+\alpha)^{-1}$ we can write
% \begin{equation}
% 	(s+\alpha)V(s) = G(s).
% \end{equation}
To arrive at the final form of the model we shall state the synaptic response kernel as a Green's function
\begin{equation}\label{GreensFuncDef}
	Lh\left( t \right) = \delta \left( t \right),
\end{equation}
where $L=\frac{d}{dt} + \alpha$ is a temporal differential operator and $\delta(t)$ is the Dirac-delta function giving 
\begin{equation}\label{FinalForm1}
	\frac{dv\left( r,t \right)}{dt} + \alpha v\left( r,t \right) = \int_\Omega  {w\left( r,r' \right)f\left( {v\left( r',t \right)} \right)dr'} + I\left(r,t\right).
\end{equation}

\section{Generating Data}
To generate iEEG/LFP data we proceeded with the assumption of spatial homogeneity for the connectivity and activation function parameters, implying $w(r,r') = w\left(|r-r'|\right)$, $\nu_0(r) = \nu_0$ and $v_0(r) = v_0$. In addition we assume the synaptic transmission delay is negligible. The model was discretized in time  (see Appendix~\ref{Time Discretization}) and a stochastic term was added to drive the system and account for uncertainty in the model giving
\begin{equation}\label{DiscreteTimeModel}
	v_{t+1}\left(r\right) = \lambda v_t\left(r\right) + \delta \int_\Omega { w\left(|r-r'|\right) f\left(v_t\left(r'\right)\right) dr'} + \delta I_t\left(r\right) + e_t\left(r\right)
\end{equation}
where $\delta$ is the time step, $\lambda = 1-\alpha\delta$ and $e_t(r) \sim \mathcal{GP}(0,\Sigma)$. The IDE formulation in Eq.~\ref{DiscreteTimeModel} was solved using a numerical method described in Appendix~\ref{Space Discretization}. To generate iEEG/LFP data we used the output function 
\begin{equation}
	y_t(r_n) = \int_{\Omega}{m\left(|r_n-r'|\right)v_t\left(r'\right)dr'} + \varepsilon_t
\end{equation}
where $r_n$ defines the location of the sensors in the field where $n=1,...,N$ indexes the sensors and $\varepsilon_t \sim \mathcal{N}\left(0,\Sigma_y\right)$. The output kernel $m(|r-r'|)$ governs the geometry of pick-up range of sensors where
\begin{equation}
	m\left(|r-r'|\right) = \exp{\left(-\frac{(r-r')^\top(r-r')}{\sigma_m^2}\right)}.
\end{equation}
